---
title: "Non-parametric estimates of basic survival quantities"
runningheader: "Tufte Handout with R Markdown" # only for pdf output
subtitle: "A brief description of the methods employed and packages used  in analysing data from a reliability use case" # only for html output
author: "Shishir Rao"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

# Introduction

In this article, we will analyse vehicle shock absorb-er failure time data and estimate a few basic survival quantities^[Failure time data is also known as survival data, life data, event-time data or reliability data, depending on the field of study.]. The data contains failure times (in kilometers driven) and the mode of failure. We will ignore the mode of failure for this article and will only consider whether a failure occurred or not i.e. censored. 

Survival function $S(t)$ is defined as the probability that a unit's time to failure ($T$) is greater than some time ($t$)^[Time is measured in kilometers in this example]. 

$$S(t) = Pr(T > t)$$

# Kaplan Meir Curve

A non-parametric estimate of the survival function is given by the product-limit estimator, proposed by Kaplan and Meir (1958)^[Kaplan, E. L. and Meier, P. Nonparametric Estimation from Incomplete Observations. *Journal of the American Statistical Association* 53 (1958): 457–481.]

$$S(t) = $$

A short description of these basic survival quantities along with a few mathematical formulas will also be inc will also be mentioned briefly, but   estimating basic survival quantities like the survival function, its standard error, confidence intervals, confidence bands and mean survival for data from a 




To see the R Markdown source of this example document, you may follow [this link to Github](https://github.com/rstudio/tufte/raw/main/inst/rmarkdown/templates/tufte_html/skeleton/skeleton.Rmd), use the wizard in RStudio IDE (`File -> New File -> R Markdown -> From Template`), or open the Rmd file in the package:

```{r eval=FALSE}
file.edit(
  tufte:::template_resources(
    'tufte_html', '..', 'skeleton', 'skeleton.Rmd'
  )
)
```

This document is also available in [Chinese](https://rstudio.github.io/tufte/cn/), and its `envisioned` style can be found [here](https://rstudio.github.io/tufte/envisioned/).

```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```
