---
title: "Non-parametric estimates of basic survival quantities"
runningheader: "Tufte Handout with R Markdown" # only for pdf output
subtitle: "A brief description of the methods employed and packages used  in analysing data from a reliability use case" # only for html output
author: "Shishir Rao"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)

library(KMsurv)
library(survival)
library(survMisc) ##Confidence bands not in survival package. survMisc is an extension of the survival package and contains confidence bands
library(tidyverse)
library(readxl)
library(readr)
library(survminer)

```

# Introduction

Recently, I have been slow-reading the book Survival Analysis: Techniques for Censored and Truncated Data, Second Edition (John P. Klein and Melvin L. Moescheberger). Although the techniques presented in this book focus on applications in biology and medicine, the same statistical tools can also be applied to disciplines ranging from engineering to economics and demography. I have a background in mechanical engineering and this project is my attempt at applying these statistical methods to data from Reliability engineering. This article is the first of many articles that I intend to write as I complete reading chapters from the book. The data sets that I plan to analyse are the ones that have been used as examples in another excellent book: Statistical Methods for Reliability Data, Second Edition (William Q. Meeker, Luis A. Escobar, Francis G. Pascual).

In this article, we will analyse vehicle shock absorb-er failure time data^[Failure time data is also known as survival data, life data, event-time data or reliability data, depending on the field of study.] and estimate a few basic survival quantities. The data contains failure times (in kilometers driven) and the mode of failure. We will ignore the mode of failure for now and will only consider whether a failure occurred or not i.e. censored. In a future article, I plan to use the different failure modes to discuss competing risks from a survival analysis context.

# Kaplan Meir Curve

Before we dive into the data and the packages used for the analysis in R, here is a brief introduction to a few basic survival concepts and their mathematical expressions. You don't strictly need to know these formulas to conduct an analysis using packages in R. I personally prefer to peek a little "under the hood", when possible, to understand what is going on in the calculations.

Survival function $S(t)$ is defined as the probability that a unit's time to failure ($T$) is greater than some time ($t$)^[Time is measured in kilometers in this example]. 

$$S(t) = Pr(T > t)$$

A non-parametric estimate of the survival function is given by the product-limit estimator, proposed by Kaplan and Meir (1958)^[Kaplan, E. L. and Meier, P. Nonparametric Estimation from Incomplete Observations. *Journal of the American Statistical Association* 53 (1958): 457–481.]. It is given by:

$$\hat{S}(t) =1  \quad \text{ if }t< t_{1}\\$$
and

$$\hat{S}(t) =\prod_{t_{i} \le t}[1 - \frac{d_{i}}{Y_{i}}]  \quad \text{ if }t_{1}\le t\\$$
where $\hat{S}(t)$ is an estimate of the survival function $S(t)$, $d_{i}$ is the number of events at time $t_{i}$ and $Y_{i}$ is the number at risk of experiencing the event at time $t_{i}$. 

The variance of the estimate of the survival function is given by Greenwood's formula:

$$\hat{V}[\hat{S}(t)] = \hat{S}(t)^2 \sum_{t_{i} \le t} \frac{d_{i}}{Y_{i}(Y_{i} - d_{i})} $$
The hazard function^[It is often called failure rate in Reliability] is another basic quantity of interest in such studies, and will be introduced in a later article. 

# Motivating example: Shock Absorber Data

In this example, we are interested in finding a non-parametric estimate of the survival function in order to gauge the probability that a shock absorber will survive a certain distance before failing.

```{r echo=TRUE, message=FALSE, warning=FALSE}
ShockAbsorber <- read_csv("ShockAbsorber.csv")

knitr::kable((ShockAbsorber), caption = "Table 1. Shock Absorber Data", align = rep('c', 3))

```

The two different modes of failures in this example will be ignored for now and we will only consider whether a shock absorber failed or not. This is indicated by the "censoring indicator" column in the above table. All the censored observations in the above table are considered to be "right" censored^[Other types of censoring are "left" and "interval" censoring.]. This means that failure has not yet occurred and would be expected to occur some time in the future (in terms of kilometers) had we continued to collect data. The concept of censoring differentiates survival analysis from other types of analysis in statistics. Information contained in the statement "the shock absorber survived at least 6950 km" is different to the information in the statement "the shock absorber failed at 6950 km". Taking censoring into account allows us to incorporate the correct information in our analysis. 

The *survival* package is the cornerstone of the survival analysis ecosystem in R and we will use the same for our analysis. The author of the package, Dr. Terry Therneau, started work on this package in 1985 and the package has evolved ever since. There are other packages that build on the *survival* package for functions that are not available within this package. For example, we will use the *survMisc* package later in this article when we build confidence bands for the survival function. 
Failure is coded as "1" and censored is coded as "0" as per convention. The *survival* package also follows this convention and will automatically detect a "1" as an event (failure).

```{r}
ShockAbsorber <- ShockAbsorber %>%
  dplyr::mutate(Event = case_when(
    `Censoring Indicator` == "Failed" ~ 1,
    `Censoring Indicator` == "Censored" ~ 0
  ))
```

Next, we build a simple Kaplan Meir survival curve using the *survfit*^[Ignore the arguments stype, ctype and conf.type for now. We will discuss them some other time] function from the *survival* package.

```{r}
fit_1 <- survival::survfit(
  survival::Surv(
    time = ShockAbsorber$Kilometers, 
    event = ShockAbsorber$Event) ~ 1 #Right hand formula for a single curve
  , stype = 1 #Survival type 1 means direct estimate of survival curve. Type 2 would have been exp(-H(t))
  , ctype = 1 #Cumulative Hazard type 1 is Nelson-Aalen. Type 2 is Fleming-Harrington correction for tied events.
  , conf.type = "log-log"
  , conf.int = 0.9
  , data = ShockAbsorber) 

#plot(fit_1, xlab = "Kilometers", ylab = "Survival Probability")

survminer::ggsurvplot(fit_1,
                      risk.table = TRUE,
                      title = "Kaplan-Meir curve",
                      xlab = "Kilometers")
```

The Kaplan Meir curve above is a step function, with steps at every *failure* time (distance in this case). There is no step when we have a censored observation, which are indicated by the short "|" marks on the curve.

The *summary* function gives us the survival probability, standard errors^[Square root of the variance formula above] and the confidence intervals at all the *failure* times.

```{r}
summary(fit_1)
```

The argument *times* in the *summary* function can be used to obtain survival probability estimates at specific times, as shown below.

```{r}

summary(fit_1, times = c(7500, 19000, 26000))

```

Note that the *n.event* column in the above summary is the number of failures between two time points. For eg: there are 3 failures in the data between 19,000 km and 26,000 km.

The probability that a shock absorber will last at least 19,000 km is 78.4% with a 95% confidence interval (61.5%, 88.5%).

## Pointwise Confidence Intervals

The simplest confidence interval is the linear confidence interval, given by 

$$\hat{S}(t_{0})\pm Z_{1-\frac{\alpha}{2}}\sigma_{s}(t_{0})\hat{S}(t_{0})$$

where $\sigma_{s}^2(t) = \hat{V}[\hat{S}(t)]/\hat{S}^2(t)$ and $Z_{1-\frac{\alpha}{2}}$ is the $(1-\frac{\alpha}{2})$ percentile of the standard normal distribution.

Although the linear confidence interval has a simple form, the normal distribution assumption used in constructing this interval is seldom valid when the sample size is small. In addition, the performance of this interval near the ends i.e $\hat{S}(t)$ close to 0 or 1 can sometimes give us CI values outside the [0,1] range, which does not make sense for a probability value. Hence, transformations are used to get better confidence intervals. 

Different transformations like the *log*, *log-log*, *logit*, *arc-sine* are available in the *survival* package by using the *conf.type* argument^[Note that the confidence intervals obtained through different transformations may not be symmetric like the linear CI]. We have used the *log-log* transformation. The formulas for the confidence intervals from the different transformations can be found in any survival modeling textbook^[I did consider including the formulas here for educational purposes, but the Latex code for typing the formulas would have taken me a very long time.]. As per the text in the *survival* package documentation by Dr. Therneau: *"Which of the choices of log, log-log, or logit is "best" depends on the details of any
particular simulation study"*. We have used the "log-log" transformation for constructing our confidence intervals.

## Confidence bands

The confidence intervals obtained above are valid only at a particular point $(t_{0})$. If we want a confidence interval for the survival function over a range of values of *t*, we will need to construct confidence bands. These confidence bands ensure that the statistical uncertainty is simultaneously quantified over a range of time and not just at a single point.^[In other words, the confidence band guarantees, with a given confidence, that the survival function falls within the band for all *t* in some interval. I see this as being somewhat similar to the correction of the significance value *\alpha* for multiple comparison tests.]

Note that although the Kaplan-Meir plot previously plotted shows the confidence region and look like a band, they are pointwise confidence intervals and are not confidence bands.

We will now construct a 95% confidence band for the range 10,000 km to 20,000 km.

The package *survMisc*, contains functions to build a confidence band. This package requires a "ten" object^[*ten* stands for time, event(s) and number at risk.], which is easily obtained as follows

```{r}
ten_ShockAbsorber <- survMisc::ten(
  survival::Surv(time = ShockAbsorber$Kilometers,
                 event = ShockAbsorber$Event)
  )
```

There are a couple of different methods to obtain the confidence bands. We will build the bands based on the approach suggested by Nair (1984)^[Nair, V. N. Confidence Bands for Survival Functions with Censored Data: A Comparative Study. *Technometrics* 14 (1984): 265–275.]. These bands are call Equal Probability or *EP* bands and are proportional to the pointwise confidence intervals.

```{r warning=FALSE}
survMisc::ci(ten_ShockAbsorber,
   CI = "0.9", #As the function currently relies on lookup tables, currently only 90%, 95% (the default) and 99% are supported.
   how = "nair", #Method to create confidence band
   trans = "log", #Transformation to use. Note that "log" in the survMisc package corresponds to the log transformation in Klein-Moeschberger
   tL = 10000,
   tU = 20000)
```

The confidence band is expected to be wider than the pointwise confidence intervals. The lower and upper values of the 90%^[Note that the calculation of the confidence bands depend on lookup tables inside the survMisc package. These tables have only been constructed for the 90%, 95% and 99% confidence leves, due to which these 3 values are the only permissible values in the CI argument to the survMisc function.] confidence band at 19,000 km is (51%, 91%) whereas the point wise confidence interval at 19,000 km, which we found previously, is (61.5%, 88.5%).


## Mean Survival Time

The mean time to failure of shock absorbers can also be easily calculated using the *survival* function. Before we directly apply the function and find the mean, there is one small detail that needs to be clarified.

The mean time to failure is a function of the survival function. To be more precise, the mean time to failure is the total area under the survival curve. If the last observation is an "event" (i.e. "failure"), the Kaplan-Meir curve will have a step and the survival probability will drop to 0 at this point. But if the last observation is a censored observation, like we see in our shock absorber example, there is no step and the curve is open at this end. We do not know when the last observation experienced failure. Clearly, this will have an impact on the area under the survival curve. Hence, we are only able to calculate the "restricted" mean, where our calculation of the mean is restricted to a certain time $\tau$. Efron's tail correction suggests treating the last observation as an "event" in case it is censored. 

The mean time to failure of the shock absobers, restricting our calculation of the mean to the highest observation (28,100 km) is obtained as follows:

```{r}
print(fit_1, rmean = 28100)

```

The mean time to failure (restricted) is 22,875 km. The lower and upper interval limits shown are for the median. The confidence limits for the mean can be calculated as follows:

$$\hat{\mu}_{\tau} \pm Z_{1-\frac{\alpha}{2}} \sqrt{\hat{V}[\hat{\mu}_{\tau}]}$$
where $\sqrt{\hat{V}[\hat{\mu}_{\tau}]}$ is the standard error of the restricted mean given in the output above and $Z_{1-\frac{\alpha}{2}}$ is the $(1-\frac{\alpha}{2})$ percentile of the standard normal distribution. The confidence interval of the mean is thus given by

```{r}
22875 - ((1283)*qnorm(0.95))
22875 + ((1283)*qnorm(0.95))
```

The restricted mean time to failure is 22,875 km with a 90% confidence interval (20764.65, 24985.35)

```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```
